Rewind: Automatically Generated\\Digital Memories from Mobile Geolocation Data

### 2016-09-21 paper/uist_2016/paper.tex
memory

\subsection{Within-subject study}
\subsubsection{Semantic Memory} 
Within-subject comparison in our study reveals that participants are significantly better at recalling semantic memory of a particular location with \textit{Rewind} videos than with static street-view images, $\chi^2(1, N=356) = 96.987$, $p <.01$. When being presented with a street-side image retrieved based on their location data, participants were able to recognize places visited 50\% of the time; yet recognition rate of locations reached 97\% when participants started playing accompanied \textit{Rewind} video, with a significant increase of 47\%. It is perhaps not too surprising as cinematic media like \textit{Rewind} video would provide users with more contextual meanings and retrival cues for semantic processing, as three participants (P4, P6 and P10) have once expressed their contentment after watching \textit{Rewind} videos by claiming ``I recognize the photo now!'' when their initial attempts with static images recognition were fruitless.

%
\subsubsection{Episodic Memory}
It could possibly be inferred that participants are likely to benefit more from \textit{Rewind} videos than static street-view images at recalling episodic memory. Although participants were not explicitly asked about trip rememberance when presented with static images as opposed to what they were asked after watchiing \textit{Rewind} videos, location rememberance from static images may serve as an upper-bound estimation of how much an individual could recall about a particular trip; since if participants fail to form any cognitive association with the locations depicted in the images, it is very unlikely that they would be able to associate target itineraries with the locations shown. Thus, it is perhaps safe to presume that location rememberance with images bounds above the maximum likelihood of trip rememberance with the same images presented to the participants, and we used location rememberance from image recognition as an upper-limit gauge against trip rememberance from \textit{Rewind} videos thereafter.

% h - add another table, only abt memorability tho
% explain distance groups in the beginning
% where is home? need to explain this part in system?
Using the aforementioned presumption, our study examined effects of \textit{Rewind} videos on episodic processing in metric-specific groups, as well as the performance overall. Our study shows that comparing to static street-view images, recollection of event-driven memories using \textit{Rewind} videos is significantly improved for trips taken that were 100 - 1,000 km (62 - 620 miles) away from home by 50\%, $\chi^2(1, N=20) = 3.215$, $p <.05$; and for trips that were least frequently taken by about 20\%, $\chi^2(1, N=161) = 5.977$, $p <.01$. Our study also suggested that GPS users would benefit more from \textit{Rewind}, even though the difference was mildly significant by an increase about 15\%, $\chi^2(1, N=118) = 2.145$, $p = .0715 (<.1)$. 
% h - prob explain this a little bit more

One plausible explanation of the improvement may be inherent to the encoding of episodic memory in the first place. As suggested by the levels-of-processing theory(ref), encoding efficacy is heavily dependent on the levels to which stimuli are being actively or passively processed. Not surprisingly, superficial mental engagement would produce weaker, and less enduring cognitive representation which would demand more contextually accurate and abundant cues for memory triggery, and vice versa. Oftentimes, the cognitive representation resulted from shallow processing may not even be retrievable at all as it is simply not available for retrieval.  As previously described, both of the trips revealed in our results, those that were taken farthest away from home and those that were least frequently taken, may not be special enough to engage deep memory encoding nor often enough to reinforce incidental remembrance. This means that it would be more difficult for participants to recall these trips if the provided cues are not effective enough. Since \textit{Rewind} videos provide richer and more contextually meaningful cues than static street-side images, it is reasonable that \textit{Rewind} significantly outperforms plain images as suggested by our statistical analysis. While being asked about location and trip rememberance on one of his least frequently taken routes, P7 was not able to identify the location from the static image at first; yet he recognized the trip from \textit{Rewind} video as ``one street (in the video) helps me (him) remember". 
%The Effects of Divided Attention on Encoding and Retrieval Processes in Human Memory 
% http://www2.psych.ubc.ca/~pgraf/Psy583Readings/Craik%20et%20al%201996.pdf
% Craik et al. 1996

The extent to which attention is dedicated while taking a trip may be another possible explanation for the observed improvement using Rewinds. This is perhaps particularly applicable to explain why participants who were self-identified regular GPS users tend to benefit more from Rewinds with episodic memory recalling. Laboratory work done by Craik et al. 1996 (ref) has demonstrated that the encoding of memory is more effective under full-attention conditions than divided-attention conditions.  This means that regular GPS users, who are often directed step-by-step on how to navigate through their environments, may be less likely to engage their attention while taking their routes, leading to less available cognitive representation for retrieval. P3, a self-identified regular GPS user who had 30\% recalling rate using static images and 80\% trip remembrance rate using Rewinds said ``I really don't pay attention to my surroundings when I'm driving. (I am) Focused (focused) on GPS and road"  Under this circumstance, it is almost certain that Rewinds would provide a more successful access to episodic memory as it provides more spatial-temporal cues for memory retrieval, by not only adjusting image attributes to reflect timely and weather changes, but also mimicing the motions of the users when they travel down a particular path.

It may appear to be puzzling that participants seem to recall less about specifics of their trips from \textit{Rewind} videos compared to static images when they traveled alone, with a significant decrease around 40\%, $\chi^2(1, N=168) = 33.831$, $p <.01$; when the trips were most frequently taken, with a significant decrease of 20\%, $\chi^2(1, N=161) = 5.977$, $p <.01$; and when the trips taken were 10 - 100 meters (33 - 328 feet) away from home, with a mildly significant decrease of 50\%, $\chi^2(1, N=16) = 2.25$, $p <.1$.

It could possibly be inferred that participants are likely to benefit more from \textit{Rewind} videos than static street-view images at recalling episodic memory. Although participants were not explicitly asked about trip remembrance when presented with static images, location rememberance from static images may serve as an upper-bound estimation of how much an individual could recall about a particular trip. %- tbc

>>>>

\subsection{Within-subject study}
\subsubsection{Semantic Memory} 
Within-subject comparison in our study reveals that participants are significantly better at recalling semantic memory of a particular location with \textit{Rewind} videos than with static street-view images, $\chi^2(1, N=356) = 96.987$, $p <.01$. When being presented with a street-side image retrieved based on their location data, participants were able to recognize places visited 50\% of the time; yet recognition rate of locations reached 97\% when participants started playing accompanied Rewinds, with a significant increase of 47\%. It is perhaps not too surprising as cinematic media like Rewinds would provide users with more contextual meanings and retrival cues for semantic processing, as three participants (P4, P6 and P10) have once expressed their contentment after watching Rewinds by claiming ``I recognize the photo now!'' when their initial attempts with static images recognition were fruitless.

%
\subsubsection{Episodic Memory}
It could possibly be inferred that participants are likely to benefit more from Rewinds than static street-view images at recalling episodic memory. Although participants were not explicitly asked about trip rememberance when presented with static images as opposed to what they were asked after watchiing Rewinds, location rememberance from static images may serve as an upper-bound estimation of how much an individual could recall about a particular trip; since if participants fail to form any cognitive association with the locations depicted in the images, it is very unlikely that they would be able to associate target itineraries with the locations shown. Thus, it is perhaps safe to presume that location rememberance with images bounds above the maximum likelihood of trip rememberance with the same images presented to the participants, and we used location rememberance from image recognition as an upper-limit gauge against trip rememberance from Rewinds thereafter.

% h - add another table, only abt memorability tho
% explain distance groups in the beginning
% where is home? need to explain this part in system?
Using the aforementioned presumption, our study examined effects of Rewinds videos on episodic processing in metric-specific groups, as well as the performance overall. Our study shows that comparing to static street-view images, recollection of event-driven memories using \textit{Rewind} videos is significantly improved for trips taken that were 100 - 1,000 km (62 - 620 miles) away from home by 50\%, $\chi^2(1, N=20) = 3.215$, $p <.05$; and for trips that were least frequently taken by about 20\%, $\chi^2(1, N=161) = 5.977$, $p <.01$. Our study also suggested that GPS users would benefit more from Rewinds, even though the difference was mildly significant by an increase about 15\%, $\chi^2(1, N=118) = 2.145$, $p = .0715 (<.1)$. 
% h - prob explain this a little bit more

One plausible explanation of the improvement may be inherent to the encoding of episodic memory in the first place. As suggested by the levels-of-processing theory(ref), encoding efficacy is heavily dependent on the levels to which stimuli are being actively or passively processed. Not surprisingly, superficial mental engagement would produce weaker, and less enduring cognitive representation which would demand more contextually accurate and abundant cues for memory triggery, and vice versa. Oftentimes, the cognitive representation resulted from shallow processing may not even be retrievable at all as it is simply not available for retrieval.  As previously described, both of the trips revealed in our results, those that were taken farthest away from home and those that were least frequently taken, may not be special enough to engage deep memory encoding nor often enough to reinforce incidental remembrance. This means that it would be more difficult for participants to recall these trips if the provided cues are not effective enough. Since Rewinds provide richer and more contextually meaningful cues than static street-side images, it is reasonable that Rewinds significantly outperforms plain images as suggested by our statistical analysis. While being asked about location and trip rememberance on one of his least frequently taken routes, P7 was not able to identify the location from the static image at first; yet he recognized the trip from Rewinds as ``one street (in the video) helps me (him) remember". 
% The Effects of Divided Attention on Encoding and Retrieval Processes in Human Memory 
% http://www2.psych.ubc.ca/~pgraf/Psy583Readings/Craik%20et%20al%201996.pdf
% Craik et al. 1996

The extent to which attention is dedicated while taking a trip may be another possible explanation for the observed improvement using Rewinds. This is perhaps particularly applicable to explain why participants who were self-identified regular GPS users tend to benefit more from Rewinds with episodic memory recalling. Laboratory work done by Craik et al. 1996 (ref) has demonstrated that the encoding of memory is more effective under full-attention conditions than divided-attention conditions.  This means that regular GPS users, who are often directed step-by-step on how to navigate through their environments, may be less likely to engage their attention while taking their routes, leading to less available cognitive representation for retrieval. P3, a self-identified regular GPS user who had 30\% recalling rate using static images and 80\% trip remembrance rate using Rewinds said ``I really don't pay attention to my surroundings when I'm driving. (I am) Focused (focused) on GPS and road"  Under this circumstance, it is almost certain that Rewinds would provide a more successful access to episodic memory as it provides more spatial-temporal cues for memory retrieval, not only by adjusting image attributes to reflect timely and weather changes, but also by mimicing the motions of the users\' when they travel down a particular path.

It may appear to be puzzling that participants seem to recall less about specifics of their trips from Rewinds compared to static imageries when they traveled alone, with a significant decrease around 40\%, $\chi^2(1, N=168) = 33.831$, $p <.01$; when the trips were most frequently taken, with a significant decrease of 20\%, $\chi^2(1, N=161) = 5.977$, $p <.01$; and when the trips taken were 10 - 100 meters (33 - 328 feet) away from home, with a mildly significant decrease of 50\%, $\chi^2(1, N=16) = 2.25$, $p <.1$. One possible reason for this would be it is hard for participants to tie a repeatedly taken trip with a specific instance in time, and trips taken near home could also be the ones that are most frequently taken. However, since our study was using location rememberance from images as an upper-limit estimation about participants' trip rememberance from the images, it is also likely that the actual trip rememberance from images could be lower.

Overall, trip rememberance from Rewinds is about 4\% higher than that from static street-view images. Although chi-squared one-tailed test did not reveal this difference to be statistically significant, $\chi^2(1, N=356) = 0.41$, $p >.1$, we are positive that the actual difference would be more significant since the trip rememberance from images was a best possible estimation from location rememberance.
